---
title: "lab4"
output: html_document
author: "Anastasia Khudoyarova"
date: "11/9/2020"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Data preparation
```{r}
df <- get(load("./income_elec_state.Rdata"))

head(df, n = 10)
```


```{r}
colnames(df)
df <- cbind(df,state=rownames(df))
rownames(df) <- NULL
names(df)[names(df) == "income"] <- "mean_household_income"
names(df)[names(df) == "elec"] <- "mean_electricity_usage"
head(df, n = 10)
```

# View
```{r}
library(ggplot2)
g <- ggplot(data=df, mapping = aes(x=mean_household_income, y=mean_electricity_usage))
g + geom_point()
```

# Clusterization
```{r}
clusters <- kmeans(df[, 1:2], centers=10, nstart = 300, iter.max=100)
```

```{r}
library(ggplot2)
g <- ggplot()
g1 <- g + 
  geom_point(df, mapping = aes(x=mean_household_income, y=mean_electricity_usage, colour=factor(clusters$cluster)), shape=1) + 
  geom_point(mapping = aes(x = clusters$center[, 'mean_household_income'], 
                                  y = clusters$center[, 'mean_electricity_usage'],
                                  size=1), shape=8)
plot(g1)
```

# Estimate number of clusters. 
## Elbow method
```{r}
kmean_withinss <- function(k) {
    cluster <- kmeans(df[,1:2], k, nstart=300, iter.max = 100)
    return (cluster$tot.withinss)
}


max_cluster <-20 
wss <- sapply(2:max_cluster, kmean_withinss)
elbow <-data.frame(2:max_cluster, wss)
```

## Plot the graph for estimation
```{r}
ggplot(elbow, aes(x = X2.max_cluster, y = wss)) +
    geom_point() +
    geom_line() +
    scale_x_continuous(breaks = seq(1, 20, by = 1)) + labs(title = "Elbow method\n", x = "number of clusters", y = "Total within-clusters sum of square") 
```

I think `k = 4` or `k = 5` is optimal variant

# Log scaled data and reevaluate
## Scale log10
```{r}
df$income_log10 <- log10(df$mean_household_income)
df$electricity_log10 <- log10(df$mean_electricity_usage)
head(df, n = 10)
```

## Clusterization for log scaled data
```{r}
clusters <- kmeans(df[, 4:5], centers=5, nstart = 300, iter.max=100)
centers <- clusters$centers
```

## Visualization
### Log10
```{r}
library(ggplot2)
g <- ggplot()
g1 <- g + 
  geom_point(df, mapping = aes(income_log10, electricity_log10, colour=factor(clusters$cluster))) 
plot(g1+geom_text(df, mapping=aes(income_log10, electricity_log10, label=state)))
```

### Simple
```{r}
g <- ggplot()
g1 <- g + 
  geom_point(df, mapping = aes(mean_household_income, mean_electricity_usage, colour=factor(clusters$cluster))) 
plot(g1)
```

They became more believable and correct view information.

# Reevaluate k for log scaled data
```{r}
kmean_withinss <- function(k) {
    cluster <- kmeans(df[,4:5], k, nstart=300, iter.max = 100)
    return (cluster$tot.withinss)
}
max_cluster <-20 
wss <- sapply(2:max_cluster, kmean_withinss)
elbow <-data.frame(2:max_cluster, wss)
```

## Plot the graph for estimation
```{r}
ggplot(elbow, aes(x = X2.max_cluster, y = wss)) +
    geom_point() +
    geom_line() +
    scale_x_continuous(breaks = seq(1, 20, by = 1)) + labs(title = "Elbow method\n", x = "number of clusters", y = "Total within-clusters sum of square") 
```

`k=5` is correct amount of clusters because we can see at the graph 

# Outlier  

As we can see outlier is Puerto Rico

### Remove outlier and get new k
```{r}
df_clear <- df[df$state != 'PR', ]
```

## Reevaluate k for df without outlier
```{r}
kmean_withinss <- function(k) {
    cluster <- kmeans(df_clear[,4:5], k, nstart=100, iter.max = 100)
    return (cluster$tot.withinss)
}

max_cluster <-20 
wss <- sapply(2:max_cluster, kmean_withinss)
elbow <-data.frame(2:max_cluster, wss)
```

## Plot for non outlier data
```{r}
ggplot(elbow, aes(x = X2.max_cluster, y = wss)) +
    geom_point() +
    geom_line() +
    scale_x_continuous(breaks = seq(1, 20, by = 1)) + labs(title = "Elbow method\n", x = "number of clusters", y = "Total within-clusters sum of square") 
```

When delete outlier we see that elbow when `k=4`

# Map

```{r}
rownames(df) <- df$state
```

```{r}
map_clusters <- kmeans(df[df$state != 'AK' & df$state != 'HI' & df$state != 'PR' & df$state != 'VI' & df$state != 'AS',4:5], centers=4, nstart = 100, iter.max=100)
map_clusters
maps::map(database = 'state', col = map_clusters$cluster, fill = TRUE)
```

# Hierarchical clustering
## Using different linkage metrices for hierarchical clustering 
```{r}
d <- dist(df[,1:2])

average= hclust(d, method = "average" )
plot(average, cex = 0.7, hang = -1, main = "Average linkage")
rect.hclust(average, k=5)

average_tree <- cutree(average, k = 5)
f <- ggplot() + 
  geom_point(df, mapping = aes(mean_household_income, mean_electricity_usage, colour=factor(average_tree))) 
plot(f + ggtitle("Hierarchical clastering. Average linkage"))
```

```{r}
single = hclust(d, method = "single" )
plot(single, cex = 0.7, hang = -1, main = "Single linkage")
rect.hclust(single, k=5)

single_tree <- cutree(single, k = 5)
f <- ggplot() + 
  geom_point(df, mapping = aes(mean_household_income, mean_electricity_usage, colour=factor(single_tree))) 
plot(f + ggtitle("Hierarchical clastering. Single linkage"))
```

```{r}
hcd <- as.dendrogram(hclust(d, method = "ward.D2" ))
nodePar <- list(lab.cex = 0.7, pch = c(NA, 19), cex = 0.7, col = "blue")
plot(hcd,  xlab = "Height", nodePar = nodePar, horiz = TRUE,
     edgePar = list(col = 2:3, lwd = 2:1), main = "Ward linkage")
```

```{r}
res <- hclust(d, method = "complete" )
grp <- cutree(res, k = 5)
plot(res, cex = 0.7, main = "Complete linkage")
rect.hclust(res, k = 5, border = 2:5)

f <- ggplot() + 
  geom_point(df, mapping = aes(mean_household_income, mean_electricity_usage, colour=factor(grp))) 
plot(f + ggtitle("Hierarchical clastering. Complete linkage"))
```


As we can see Puerto Riko is an outlier

## Log10 scaled data
```{r}
d <- dist(df[,4:5])
res <- hclust(d, method = "complete" )
grp <- cutree(res, k = 5)
plot(res, cex = 0.7, main = "Complete linkage")
rect.hclust(res, k = 5, border = 2:5)
```

Puerto Riko definitely separated from others

# Estimate amount of clusters

```{r}
library(cluster)
library(factoextra)
```

## Elbow method
```{r}
fviz_nbclust(df[,4:5], hcut, method = "wss")
```

Wss method gives `k=5`.

## silhouette method
```{r}
fviz_nbclust(df[,4:5], hcut, method = "silhouette")
```

A high average silhouette width indicates a good clustering at `k=9`

## Gap statistic
```{r}
gap_stat <- clusGap(df[,4:5], FUN = hcut, nstart = 200,
                    K.max = 20, B = 50)
fviz_gap_stat(gap_stat)
```

clusGap gives `k=1`.

## Graph for k = 5
```{r}
fviz_cluster(list(cluster = grp, data = df[,4:5]))
```

## Graph for k = 1
```{r}
grp <- cutree(res, k = 1)
fviz_cluster(list(cluster=grp, data = df[,4:5]))
```

## Graph for k = 9
```{r}
grp <- cutree(res, k = 9)
fviz_cluster(list(cluster=grp, data = df[,4:5]))
```

The most suitable variant is `k=5`: if `k=1` then all data in one cluster that is bad, if `k=9` then there is other outlier that is can be inside cluster.


# Map (for test)
```{r}
map_order <- c('AL', 'AZ', 'AR', 'CA', 'CO', 'CT', 'DE', 'DC', 'FL',
'GA', 'ID', 'IL', 'IN', 'IA', 'KS', 'KY', 'LA', 'ME',
'MD', 'MA', 'MA', 'MA', 'MI', 'MI', 'MN', 'MS', 'MO',
'MT', 'NE', 'NV', 'NH', 'NJ', 'NM', 'NY', 'NY', 'NY',
'NY', 'NC', 'NC', 'NC', 'ND', 'OH', 'OK', 'OR', 'PA',
'RI', 'SC', 'SD', 'TN', 'TX', 'UT', 'VT', 'VA', 'VA',
'VA', 'WA', 'WA', 'WA', 'WA', 'WA', 'WV', 'WI', 'WY')

rownames(df) <- df$state

clusters <- kmeans(df[map_order, 4:5], 4, nstart = 100, iter.max=100)

map_color <- clusters$cluster[map_order]

library(maps)
maps::map('state', col = map_color, fill = TRUE)
```

```{r}
elem <- factor(clusters$cluster)
levels(elem)
```